---
title: "format_data"
author: "Dan Weinberger"
date: "5/25/2021"
output:
  pdf_document: default
  html_document: default
---

## INSTRUCTIONS TO GET SURVEY DATA

1.  Go to Yale qualtrics site
2.  Navigate to Intake questionaire.
3.  Download data as .xlsx
4.  Repeat for fortnightly survey

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readxl)
library(reshape2)
library(dplyr)
library(stringr)

library(pheatmap)
library(RColorBrewer)
library(cowplot)
library(gridExtra)
library(ggplotify)
library(table1)
library(lubridate)
library(patchwork)
library(msm)
library(gee)
source('heat.fun.R')
source('heat.fun.contact.R')

```

ID 6.1 Saliva was very low under 100 ul.

ID 5.6 qpcr LytA -piaB was ran in plate A155_scope_2021_06_02 .The issue was the label: FIXED 5.5 to 5.6

ID 62.2,63.2 samples were picked up 1 month after .We didn't see any growth after incubation 37 degree over night

ID 20.5 - 21.5 to verify


What to do if we have borderline positive then negative test:

1) If weak positive for piaB (e.g., 35 Ct or higher) AND at least one of the repeats is \<40, then it is positive

2) If weak positive for piaB (e.g., 35 Ct or higher), AND (2 repeats negative) then NEGATIVE

3) if weak positive for piaB and then negative on 1 repeat performed and negative, then negative

4) If posivi, take ave of Ct among those \<40


```{r, echo=F}
 key1a <- read.csv('./Data/hh_id_key.csv')

master1 <- read_excel('./Data/confidential/SCOPE_Master_20210108.xlsx', sheet='COM SAL')


key2 <- read_excel('./Data/confidential/SCOPE Participant MRN and ID.xlsx', skip=1)
key3 <- cbind.data.frame('id'=c(100:101) , 'Household'=c(51,51) )

key4a <- bind_rows(key1a,key3)

key4a$id <- paste0('S1_', key4a$id)
key4a$Household <- paste0('S1_', key4a$Household) #SEASON 1 key


file_list <- list.files(path="./Data/Final Scope Results Drift Correction")
  
 all.res.season1 <-
   lapply(file_list, function(x){
     print(x)
     ds <- read_excel(paste0("./Data/Final Scope Results Drift Correction/",x), skip=19)
     ds$expt <- substr(x,1,4)
     ds$Sample <- as.character(ds$Sample)
     ds$Target <- as.character(ds$Target)
     ds$Cq <- as.numeric(as.character(ds$Cq))
     cols.keep <- 
       c('Cq','Sample','Target','Content','expt','Biological Set Name')
     ds.names <- names(ds)
     ds.names <- intersect(ds.names,cols.keep)
    ds <- ds[,ds.names]
      ds$Sample <- paste0('S1_', ds$Sample)

     return(ds)
   })
 
 file_list_s2 <- list.files(path="./Data/cleaned_season2")

  all.res.season2 <-
   lapply(file_list_s2, function(x){
     print(x)
     ds <- read_excel(paste0("./Data/cleaned_season2/",x), skip=19)
     ds$expt <- substr(x,1,4)
     ds$Sample <- as.character(ds$Sample)
     ds$Target <- as.character(ds$Target)
     ds$Cq <- as.numeric(as.character(ds$Cq))
     cols.keep <- 
       c('Cq','Sample','Target','Content','expt','Biological Set Name')
     ds.names <- names(ds)
     ds.names <- intersect(ds.names,cols.keep)
    ds <- ds[,ds.names]
    ds$`Biological Set Name` <- NA
    ds$Sample <- paste0('S2_', ds$Sample)
     return(ds)
   })
 
all.res <- c( all.res.season1,all.res.season2)
  
all.res <- lapply(1:length(all.res), function(x){ all.res[[x]]$plate= x 
return(all.res[[x]])
}) 

test1 <- lapply(all.res, function(x)  is.character(x$Target))

b1 <- bind_rows(all.res) 

b1$Cq[is.na(b1$Cq)] <- 45

b1$Target[is.na(b1$Target)] <- 
  b1$`Biological Set Name`[is.na(b1$Target)]

b1 <- b1[b1$Content!='Std',]

b1$Target <- gsub(' ' ,'', b1$Target, fixed=T)

b1$Target <- tolower(b1$Target)

b1$Target[grep('lyta',b1$Target)] <- 'lyta'

b1$Target[grep('piab',b1$Target)] <- 'piab'


#b1$Sample[is.na(b1$Sample)] <- b1$`sample id`[is.na(b1$Sample)]

b1 <- b1[!is.na(b1$Sample),]

b1 <- b1[-grep('neg',b1$Sample),]
b1 <- b1[-grep('NA',b1$Sample),]
b1 <- b1[-grep('NTC',b1$Sample),]
b1 <- b1[-grep('NEG',b1$Sample),]
b1 <- b1[-grep('NEC',b1$Sample),]
b1 <- b1[-grep('nec',b1$Sample),]
b1 <- b1[-grep('ntc',b1$Sample),]
b1 <- b1[-grep('apex',b1$Sample),]

b1 <- b1[-grep('-1',b1$Sample),]
b1 <- b1[-grep('-2',b1$Sample),]
b1 <- b1[-grep('-3',b1$Sample),]
b1 <- b1[-grep('-4',b1$Sample),]
b1 <- b1[-grep('-5',b1$Sample),]
b1 <- b1[-grep('QC',b1$Sample),]

b1$Sample <- gsub('rep','',b1$Sample)

#b1$Sample <- round(as.numeric(b1$Sample),1)

c1 <- b1[, c('Sample','Cq','Target','plate')]

c1$Cq[is.na(c1$Cq)] <- 45

c1.m <- reshape2::melt(c1, id.vars= c('Sample','Target','plate'))

##If sample tested multiple times call pos if pos on multiple runs. Ct<40=pos

c1.m <- c1.m %>%
  group_by(Sample, Target) %>%
  mutate(N_tests= n(), n_pos=sum(value<40), ct_pos=mean(value[value<40], na.rm=T), ct_min=min(value, na.rm=T) ) %>%
  ungroup() 

#if negative, set ct to 45
c1.m$ct_pos[is.nan(c1.m$ct_pos)] <- 45

c1.m$ct_pos2 <- c1.m$ct_pos

#If min Ct value<35, use ave value of positive regardless of number of tests performed.
c1.m$ct_pos2[c1.m$ct_min <= 35] <- c1.m$ct_pos[c1.m$ct_min <= 35] #if ave is <35, call positive regardless of N tests

#If min Ct>35, and 3 or more tests performed, with only 1 positive, call it a negative
c1.m$ct_pos2[c1.m$N_tests>=3 & c1.m$n_pos<=1 & c1.m$ct_min > 35] <- 45 #if 3 or more tests, and 0 or 1 are weakly positive or neg, then call it negative

#If min >35, with at least 2 tests performed and at least 2 were positive, call it positive
c1.m$ct_pos2[c1.m$N_tests>=2 & c1.m$n_pos>= 2 & c1.m$ct_min > 35] <- 
  c1.m$ct_pos[c1.m$N_tests>=2 & c1.m$n_pos>= 2 & c1.m$ct_min > 35] #if 2 or more tests, and 0 or 1 are weakly positive or neg, then call it negative

#if min Ct>35, N tests is <=2 and only 1 positive, call it negative
c1.m$ct_pos2[c1.m$N_tests<=2 & c1.m$n_pos< c1.m$N_tests] <- 45 #if 2 or fewer tests, and at least 1 is positive, call positive




c1.c <- reshape2::dcast(c1.m[,c('Sample','variable','Target','ct_pos2')], Sample +variable ~ Target, fun.aggregate = min, fill=9999, value.var='ct_pos2')

c1.c$pos <- c1.c$lyta<40 & c1.c$piab<40

write.csv(c1.c, './Data/cleaned_file_s2.csv')


all.ids <- unique(c1.c$Sample)
all.ids <- all.ids[substr(all.ids,1,2)=='S2']
all.ids <- gsub(".*_","",all.ids)
all.ids <- sub('\\..*', '', all.ids)
all.ids <- sort(as.numeric(as.character(unique(all.ids))))

key4b <- cbind.data.frame('id'=all.ids ,'Household'=rep(1:(length(all.ids)/2),each=2) )      
key4b$id <- paste0('S2_', key4b$id)
key4b$Household <- paste0('S2_', key4b$Household) #SEASON 1 key

key1ab <- bind_rows(key4a, key4b)

```

```{r, fig.width=5, fig.height=5}
plot(c1.c$lyta, c1.c$piab, xlim= c(45,0), ylim=c(45,0), xlab='lytA', ylab='piaB', bty='l')
abline(a=0, b=1)
abline(h=40, v=40, col='gray', lty=2)
```

```{r, echo=F}

c1.m$ID <- sub("\\.[0-9]+$", "", c1.m$Sample)

c1.m$time_1a <- sub('.*\\.', '', c1.m$Sample)
c1.m$time_2a <- as.numeric(substr(c1.m$time_1a,1,1)) 
c1.m$time_2b <- as.numeric(substr(c1.m$time_1a,2,2))
c1.m$time_2b[is.na(c1.m$time_2b)] <- 0
c1.m$time <- round(c1.m$time_2a + c1.m$time_2b/10)

c1.m <- c1.m[order(c1.m$ID, c1.m$time),]

c1.m <- merge(c1.m, key1ab, by.x='ID',by.y='id', all.x=T)

c1.m <- c1.m[!(c1.m$ID %in% c('S1_38','S1_39', 'S1_48','S1_49', 'S1_58', 'S1_59')),]

d1 <- dcast(c1.m[,c('ID','Target','time','ct_pos2')],  ID +Target ~ time, fun.aggregate = min, na.rm=T, fill=9999, value.var = 'ct_pos2')


write.csv(d1, './Data/Result1_S2.csv')

d1.alt <- dcast(c1.m[,c('ID','Target','time','ct_pos2')],  ID +time~ Target, fun.aggregate = min, na.rm=T, fill=9999, value.var = 'ct_pos2')

d1.alt <- d1.alt[!(d1.alt$ID %in% c('S1_38','S1_39', 'S1_48','S1_49', 'S1_58', 'S1_59')),] #these IDs were not included in study

write.csv(d1.alt, './Data/Result2_S2.csv') #Anne's preferred format

c1.m <- c1.m[!(c1.m$ID %in% c('S1_38','S1_39', 'S1_48','S1_49', 'S1_58', 'S1_59')),] #these IDs were not included in study

d1.a <- acast(c1.m[,c('ID','Target','time','ct_pos2','Household')], Target ~ ID +Household ~ time, fun.aggregate=min, na.rm=T, fill=9999, value.var = 'ct_pos2')

d1.ds <- reshape2::dcast(c1.m[,c('ID','Target','time','ct_pos2','Household')],   Household +ID+ time ~ Target, fun.aggregate=min, na.rm=T, fill=9999, value.var = 'ct_pos2')

d1.ds <- d1.ds %>%
  group_by(Household ,time) %>%
  arrange(ID) %>%
  mutate(HH_order=row_number() ) %>%
  ungroup()

```

## Heatmaps

piaB

```{r, fig.width=4, fig.height=8}
piab.plot <- heat.fun(target='piab')

# breakslist<-seq(15,45,by=1)
# 
# cols1 <- colorRampPalette(c('red4','red','darkorange','orange','gold','lightgoldenrod1','white','white'))(length(breakslist))
# 
# p1 <- ggplot(d1.ds, aes(time, HH_order, fill= piab)) +
#   geom_tile() +
#   scale_fill_viridis_c(option = "B", direction = -1) +
# 
#   theme_classic()+
#   ylab('Serotype tested')+
#    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
#   facet_wrap(~Household, ncol=3)+
#   theme(
#   strip.text.x = element_blank()
# )
# p1
```

lytA

```{r, fig.width=4, fig.height=8}

lyta.plot <- heat.fun(target='lyta')

```

Contacts

## Summary stats

```{r}
c1.m.m <- melt(c1.m[,c('Target','ID','Sample','ct_pos2','Household') ],
                     id.vars=c('Target','ID','Sample','Household')) 
                     
c1.m.c <- dcast(c1.m.m, ID + Household + Sample  ~Target, fun.aggregate = min, fill=999)

#N people
N.id <- length(unique(c1.m.c$ID))
N.id

#N households
N.hh<- length(unique(c1.m.c$Household))
N.hh



#N samples
nrow(c1.m.c)
#length(unique(c1.m.c$Sample[c1.m$Target=='piab']))
#length(unique(c1.m.c$Sample[c1.m$Target=='lyta']))

#N positive
sum(c1.m.c$piab<40 )

sum(c1.m.c$piab<40  &c1.m.c$lyta>=40)
sum(c1.m.c$piab<40  &c1.m.c$lyta<40)

#% positive
sum(c1.m.c$piab<40 )/nrow(c1.m.c)

#N positive lytA
sum(c1.m.c$lyta<40 )

sum(c1.m.c$lyta<40 )/nrow(c1.m.c)


#Count N people with piaB positives by IDs
c1.m.spl.id <- split(c1.m.c, c1.m.c$ID)
N_pos_people <- sum(sapply(c1.m.spl.id, function(x){
  min(x$piab) <40 
}
  ))
N_pos_people
N_pos_people/N.id

#Count N HH with piab positives by IDs
c1.m.spl.id <- split(c1.m.c, c1.m.c$Household)
n_pos_hh <- sum(sapply(c1.m.spl.id, function(x){
  min(x$piab) <40 
}
  ))
n_pos_hh
n_pos_hh/N.hh

## Ave age
age1 <- read_excel('./Data/confidential/AGE Scope Participants samples pick-up 06-15-2021.xlsx', skip=1)
names(age1)[1] <- 'ID'
age1 <- age1[ which(age1$ID %in% unique(c1.m.c$ID)),]
mean(age1$AGE)
range(age1$AGE)

```

## Merge in results with master file

--To do: -harmonize column names between fortnightly and initial questionaire so we can merge -Fix dates -Check missing visitNs

Intake questionaire

```{r, echo=F}
q0a <-  read_excel('./Data/confidential/SCOPE Intake questionnaire_September 24, 2021_11.57.xlsx', skip=1)

q0a$`Date of Visit` <- parse_date_time(q0a$`Date of Visit`, orders=c('d-b-Y','d/b/Y','d-m-Y','m/d/Y','m/d/y','Y-m-d'))

q0b <- read_excel('./Data/confidential/SCOPE Intake questionnaire_January 18, 2022_09.50.xlsx', skip=1)
names(q0b) <- names(q0a)
q0b$`Date of Visit` <- parse_date_time(q0b$`Date of Visit`, orders=c('d-b-Y','d/b/Y','d-m-Y','m/d/Y','m/d/y','Y-m-d'))


q0c <- read_excel('./Data/confidential/SCOPE Intake questionnaire_December 7, 2021_13.05.xlsx', skip=1)
names(q0c) <- names(q0a)
q0c$`Date of Visit` <- parse_date_time(q0c$`Date of Visit`, orders=c('d-b-Y','d/b/Y','d-m-Y','m/d/Y','m/d/y','Y-m-d'))

q0 <- bind_rows(q0a, q0b, q0c) 


select.colnames <- c('MRN', 
            "SCOPE ID (eg. 0007, 0044, 0072 etc) if known/listed", 
            "Date of Visit","Visit Location - Selected Choice",
            "What is your date of birth? (DD-MM-YYYY)", 
            "1.     Are you experiencing any of the following symptoms today?   - Q8#1 - Nasal congestion",
            "1.     Are you experiencing any of the following symptoms today?   - Q8#1 - Coughing",
            "1.     Are you experiencing any of the following symptoms today?   - Q8#1 - Running nose",
            "1.     Are you experiencing any of the following symptoms today?   - Q8#1 - Sore throat", 
            "1.     Are you experiencing any of the following symptoms today?   - Q8#1 - Fever", 
            "1.    \nHave you had a positive test for COVID?",
            "If\nyes, when did you have this positive test? (DD-MM-YYYY)",
            "Have\nyou received the pneumonia vaccine?"  , 
            "If yes,\nwhen did you receive this vaccine? (DD-MM-YYYY)",
           "Have\nyou been hospitalized for pneumonia in the last month (30 days)?" , 
           "Have\nyou taken any antibiotics in the last month (30 days)?",  
           "Weight",
           "Height (feet, inches)","
           Gender - Selected Choice","What is your highest level of education?", 
           "How would you describe your ethnicity? - Selected Choice"  ,
           "What is the relationship between yourself and the person you are enrolling into the study with?", 
           "How long have you been living together? - Years" ,
           "How much time do you spend together each day? - ___ hours", 
           "Have you received the flu shot this season?",
           "Have you had any of the following symptoms in the past 2 weeks? - Q39#1 - Nasal congestion" ,
          "Have you had any of the following symptoms in the past 2 weeks? - Q39#1 - Coughing",
         "Have you had any of the following symptoms in the past 2 weeks? - Q39#1 - Running nose",
         "Have you had any of the following symptoms in the past 2 weeks? - Q39#1 - Sore throat" ,
         "Have you had any of the following symptoms in the past 2 weeks? - Q39#1 - Fever",
         "Have you taken any medications for or been told by a health care provider that you have any of th... - Q40#1 - 1.\tImmunodeficiency (Examples include: HIV/AIDS, corticosteroid medication like “Prednisone”, medicine or therapy to treat a cancerous tumor or mass, have had your spleen removed or have chronic kidney disease)",
         "Have you taken any medications for or been told by a health care provider that you have any of th... - Q40#1 - 2.\tDiabetes",
         "Have you taken any medications for or been told by a health care provider that you have any of th... - Q40#1 - 5c. Chronic lung disease: Asthma",
          "Do you currently smoke cigarettes or e-cigarrettes?",
         "Do you have any regular contact with children?",
         "What is the age range of the children you have had contact with? - Selected Choice" ,
         "How often do you usually have contact with children?" ,
         "How much contact per day do you have?",
         "Time of sample collection:" 
         )
q0 <- q0[,c(which(names(q0) %in% select.colnames),grep('Are you experiencing any of the following symptoms today', names(q0)) )]

#q0b <- q0b[,c(which(names(q0b) %in% select.colnames),grep('Are you experiencing any of the following symptoms today', names(q0b)) )]

rename.intake <- c('MRN','scope_id', 'visit_date','visit_location', 'dob','pos_covid_past', 'pos_covid_date', 'pneu_vax', 'pneu_vax_date','pneumonia_hosp', 'recent_abx', 'weight','height','education','ethnicity', 'hh_relationship', 'relationship_duration','hh_contact_time','flu_shot', 'recent_nasal', 'recent_cough', 'recent_runny_nose', 'recent_sore_throat', 'recent_fever','immuno_meds','diabetes_meds','asthma_meds', 'current_smoke', 'contact_children','age_child_contacts', 'frequency_child_contacts', 'time_day_child_contacts', 'time_sample', 'current_nasal','current_cough', 'current_runny_nose', 'current_sore_throat','current_fever', 'other_symp1','othr_symp2')

names(q0) <- rename.intake

q0$visit_date[q0$visit_date %in% c( '12/APR/2021','12 Apr 2021','12/14/2021')] <- '2021-04-12'
q0$visit_date_clean <- parse_date_time(q0$visit_date, orders=c('d-b-Y','d/b/Y','d-m-Y','m/d/Y','m/d/y','Y-m-d'))
q0$visitN <- '1'

q0$MRN <- as.numeric(gsub('MR','',q0$MRN))


q0.baseline.vars <- c('dob','scope_id',"pos_covid_past","pos_covid_date","pneu_vax", "pneu_vax_date",'ethnicity','weight','height','education', "immuno_meds" ,"diabetes_meds", "asthma_meds","flu_shot","hh_relationship", "relationship_duration")
```

Fortnightly questionnaire and intake cleaning...do not run each time

```{r}
q1 <- read_excel('./Data/confidential/SCOPE Fortnightly Questionnaire_September 24, 2021_12.00_clean.xlsx', skip=1)

orig.colnames <- names(q1)

select.colnames.q1 <- c('MRN', "Fortnightly visit number:", "SCOPE ID if known/listed (eg. 0002, 0027, 0044, 0072...)", "Date of Visit", "Visit Location - Selected Choice","Have you taken part in any social activities or outings during the past two weeks?", "What sorts of activities have you participated in? - Selected Choice" , "Have you had any contact with children in the past two weeks?" , "If yes, what is the age range of the children you have had contact with? - Selected Choice","How often do you usually have contact with children?" , "How much contact per day do you have?", "Have you been tested for COVID in the last two weeks?" , "Have you had any sick visits to the doctor or been hospitalized in the last two weeks?" , "Have you received any new vaccines in the last two weeks? - Selected Choice"   ,"Time of sample collection:" ,'Have you taken any antibiotics in the last two weeks?' )

rename.q1 <-c('MRN','scope_id', 'visit_date', 'visitN', 'visit_location','recent_social_activities','social_activity_type', 'contact_children','age_child_contacts', 'frequency_child_contacts', 'time_day_child_contacts','recent_covid_test', 'recent_doctor','recent_abx', 'recent_vaccines', 'sample_time', 'current_nasal','current_cough', 'current_runny_nose', 'current_sore_throat','current_fever', 'other_symp1','othr_symp2','recent_nasal', 'recent_cough', 'recent_runny_nose', 'recent_sore_throat', 'recent_fever' , 'recent_other_symp1','recent_other_symp2')
```

```{r, eval=F}
q1 <- q1[,c(which(names(q1) %in% select.colnames.q1),grep('Which symptoms are you experiencing today', names(q1)),grep('Have you had any of the following symptoms', names(q1))  )]

names(q1) <- rename.q1

q1$visit_date_clean <- parse_date_time(q1$visit_date, orders=c('d-b-Y','d/b/Y','d-m-Y','d-m-y','m/d/Y','m/d/y','Y-m-d'))
           
q1$visit_date <- as.Date(q1$visit_date)

q1$MRN <-  gsub("mr","", q1$MRN )
q1$MRN <- as.numeric(gsub('MR','', q1$MRN)) 

q1 <- bind_rows(q0[,-which(names(q0) %in% q0.baseline.vars)], q1) #combine q0, q1

q1$MRN <-  gsub("_visit#2.pdf","", q1$MRN )
q1$MRN <-  gsub("MR","", q1$MRN )
q1$MRN <-  gsub("mr","", q1$MRN )
q1$MRN <-  as.numeric(q1$MRN)
q1$visitN <- gsub('V','', q1$visitN)
q1$visitN <- gsub('v','', q1$visitN)
q1$visitN <- as.numeric(q1$visitN)

#q1 <- merge(q1, q0[,c('MRN',q0.baseline.vars)], by='MRN', all=T)
#Output file for manual cleanring, and read back in cleaned version
write.csv(q1, './data/confidential/survey_data_to_clean.csv')
```

read in newly entered surveys

```{r}
q1b1 <- read_excel('./data/confidential/SCOPE Fortnightly Questionnaire_December 7, 2021_13.03.xlsx', skip=1)
names(q1b1) <- orig.colnames

q1b2 <- read_excel('./data/confidential/SCOPE Fortnightly Questionnaire_December 11, 2021_13.30.xlsx', skip=1)
names(q1b2) <- orig.colnames
q1b2$`Date of Visit` <- as.Date(q1b2$`Date of Visit`, '%m/%d/%Y')

q1b3 <- read_excel('./data/confidential/SCOPE Fortnightly Questionnaire_January 18, 2022_09.02.xlsx', skip=1)
names(q1b3) <- orig.colnames
q1b3$`Date of Visit` <- as.Date(q1b3$`Date of Visit`, '%m/%d/%Y')

q1b <- bind_rows(q1b1, q1b2,q1b3)




q1b <- q1b[,c(which(names(q1b) %in% select.colnames.q1),grep('Which symptoms are you experiencing today', names(q1b)),grep('Have you had any of the following symptoms', names(q1b))  )]

names(q1b) <- rename.q1

q1b$MRN <- as.numeric(q1b$MRN)

q1b$visitN <- as.numeric(q1b$visitN)
#q1b <- unique(q1b) #de-deuplicate

q1b <- q1b %>%
  group_by(MRN, visitN) %>%
  mutate(repN=row_number()) %>%
  ungroup()

q1b <- q1b[q1b$repN==1,]
```


```{r}
q2 <- read_excel('./data/confidential/survey_data_to_clean_DMW.xlsx')
q2$visit_date <- as.Date(as.numeric(q2$visit_date), origin='1900-01-01')
q2$visit_date_cleaned <- as.Date(as.numeric(q2$visit_date_cleaned), origin='1900-01-01')
q2$visit_date <- q2$visit_date_cleaned
  
q0$visitN <- as.numeric(q0$visitN)
q0$MRN <- as.numeric(q0$MRN)

q2 <- bind_rows(q2, q0[,-which(names(q0) %in% q0.baseline.vars)])
q2 <- bind_rows(q2, q1b)

q2$MRN[q2$MRN==17330176] <- 1733076 #fix typo
q2$MRN[q2$MRN==1156975 ] <- 1156974 #fix typo
q2$MRN[is.na(q2$MRN) & q2$visit_date_cleaned=='2021-03-17'] <- 4149574

q2$visit_date_cleaned <- q2$visit_date
q2$MRN_cleaned[is.na(q2$MRN_cleaned)] <-
  q2$MRN[is.na(q2$MRN_cleaned)]

##Summary of which people are missing which days
q2.m <- reshape2::melt(q2[, c('MRN_cleaned','visitN') ], id.vars=c('MRN_cleaned','visitN'))
q2.c <- reshape2::dcast(q2.m, MRN_cleaned ~visitN, fun.aggregate = length)
write.csv(q2.c,'./Data/confidential/MRN_surveys_missing.csv')
q2 <- q2[,-which(names(q2) %in% q0.baseline.vars)] #baseline vars are messed up in merge
q0$MRN <-  gsub("_visit#2.pdf","", q0$MRN )
q0$MRN <-  gsub("MR","", q0$MRN )
q0$MRN <-  gsub("mr","", q0$MRN )
q0$MRN <-  as.numeric(q0$MRN)
q2$MRN <-  gsub("MR","", q2$MRN )
merge_key <- unique(q2[, c('MRN',  "MRN_cleaned" )])
q0a <- merge(q0, merge_key, by='MRN', all=T)

q3 <- merge(q2, q0a[,which(names(q0a) %in% c('MRN_cleaned',q0.baseline.vars))], by= "MRN_cleaned", all.x=T)

q3 <- q3[q3$visitN %in% c(1,2,3,4,5,6),]

q3$visit_date_cleaned <- as.Date(q3$visit_date_cleaned)
q3$dob_clean <-  parse_date_time(q3$dob, orders=c('%m/%d/%Y', '%m/%d/%y', '%d/%m/%y','%d/%m/%y','%m-%d-%Y','%d-%m-%Y'))
q3$dob_clean[q3$dob=='1945'] <- as.Date('1945-01-01')
q3$dob_clean[q3$dob=='/1947'] <- as.Date('1947-01-01')
q3$dob_clean <- as.Date(q3$dob_clean)
q3$dob_clean <- q3$dob_clean - 100*365*(q3$dob_clean>=as.Date('2000-01-01')) #if DOB >year 2000, indicates problem with date
q3$age <- round(as.numeric(q3$visit_date_cleaned - as.Date(q3$dob_clean))/365) #only recorded at t=1
#View(q2[, c('visitN','dob_clean',"MRN..Clean.",'Visit_date..Clean.') ])
key2$mr_clean <- gsub('MR', '',key2$`MR#`,)
#This is tricky..on key 2 is minimally cleaned; should match with uncleaned MRN on q3

q4 <- merge(q3, key2, by.x="MRN_cleaned", by.y= "mr_clean" , all.x=T)

q4$flag_id <- is.na(q4$`ID Number`)
age1 <- read_excel('./Data/confidential/AGE Scope Participants samples pick-up 06-15-2021.xlsx', skip=1)

age1$MRN_cleaned <- gsub( 'MR', '',age1$`MR#`)
age1 <- age1[,c('MRN_cleaned', 'AGE' )]
age1$MRN_cleaned <- as.numeric(age1$MRN_cleaned)

q4 <- merge(q4, age1, by='MRN_cleaned', all.x=T)


##Flag duplicate survey entries

q4 <- q4 %>%
  group_by(MRN_cleaned, visitN) %>%
  mutate(repN = row_number())

q4 <- q4[q4$repN==1,]

q4 <- q4[!(q4$`ID Number`  %in% c(38,39, 48,49, 58, 59, 102, 103)),]
q4 <- q4[!(q4$`ID Number`==6 & q4$visitN==1),]

q4$ID <- paste0('S1_', q4$`ID Number`)

#View(q2[is.na(q2$age),])
#miss.dob <- q0a[q0a$dob %in% c('N/A',"19-04-2021") | is.na(q0$dob), c('dob','MRN_cleaned','MRN') ]
#write.csv(miss.dob,'./Data/confidential/check_dob.csv')
```

heat map of contacts

```{r, fig.width=4, fig.height=8}
##Summary of which people are missing which days
q4.1 <- q4

q4.1$child_contact_2_10y <- 0
q4.1$child_contact_2_10y[grep('24-59 months', q4.1$age_child_contacts)] <- 1
q4.1$child_contact_2_10y[grep('5-10 years', q4.1$age_child_contacts)] <- 1
#q4.1$child_contact_2_10y[grep('13-23 months', q4.1$age_child_contacts)] <- 1
#q4.1$child_contact_2_10y[grep('<12 months', q4.1$age_child_contacts)] <- 1

q4.1$child_contact_2_10y[is.na(q4.1$age_child_contacts)] <- NA

q4.m <- reshape2::melt(q4.1[, c('ID','visitN','child_contact_2_10y') ], id.vars=c('ID','visitN'))

q2.c <- reshape2::dcast(q4.m, `ID` ~visitN, fun.aggregate = mean, fill=9999)
names(q2.c)[1] <- 'ID'

q2.c <- q2.c[!(q2.c$ID %in% c(38,39)),] #these IDs were not included in study

q3.c <- merge(q2.c, key1ab, by.x='ID',by.y='id')
```

```{r, fig.width=4, fig.height=8}

heat.fun.contact()


```

```{r, echo=F}
hist(age1$AGE, main='Distribution of ages', xlab='Age(y)')
```

```{r, echo=F}



# master1$merge_id <- as.numeric(as.character(gsub('CS','',master1$dw_id_full)))
# 
# e1 <- merge(c1.c, master1, by.x='Sample', by.y='merge_id', all.x)
# 
# excel_date_diff <- as.numeric(as.Date("1970-01-01") - as.Date('1899-12-30')) #origin for excel and R is different
# 
# e1$`sample date`[e1$`sample date`=='2/15/25021'] <- as.character(as.numeric(as.Date('2021-02-15')) + excel_date_diff )
# 
# 
# 
# e1$pos <- 1*(e1$piab<40)
# 
# e1$sample_Date <- as.Date(as.numeric(e1$`sample date`), origin=as.Date('1899-12-30')) #use excel origin of 12-30-1899
# 
# hist(e1$sample_Date, breaks=10)
# e1$part_id <- as.numeric(e1$number)
# 
# e2 <- merge(e1, key2, by.x='part_id', by.y="ID Number", all=T)
# e2$`MR#` <- as.numeric(gsub("MR","", e2$`MR#`  ))


##NEED TO FIX THIS MERGE--MR and visitN could be missing
e3 <- merge(d1.ds, q4, by.x=c('ID','time'), by.y=c("ID",'visitN'), all=T)  


e3$social_activity_type <- as.character(e3$social_activity_type)
e3$social_activity_type[e3$recent_social_activities=='No'] <- 'No'

e3$frequency_child_contacts <- as.character(e3$frequency_child_contacts)

e3$frequency_child_contacts[e3$contact_children=='No'] <- 'No contact'

e3$piab_pos <- NA
e3$piab_pos[e3$piab >=40 & !is.na(e3$piab)] <- 0 
e3$piab_pos[e3$piab <40 & !is.na(e3$piab)] <- 1 
e3$piab_pos <- as.factor(e3$piab_pos)

e3$smoke_dic <- NA
e3$smoke_dic[grep('Yes',e3$current_smoke)] <- 1
e3$smoke_dic[grep('No',e3$current_smoke)] <- 0

e3$pneu_vax_dic <- NA
e3$pneu_vax_dic[grep('Yes',e3$pneu_vax)] <- 1
e3$pneu_vax_dic[grep('No',e3$pneu_vax)] <- 0

e3$pneu_vax_date[e3$pneu_vax_date=='N/A'] <- NA

e3$education <- tolower(e3$education)

e3$height <- gsub(' ', '', e3$height)
## export

id.ages <- unique(e3[,c('ID','AGE')])
id.ages <- id.ages[!is.na(id.ages$AGE),]

exp1 <- e3[,c('ID','time','piab', 'lyta','Household','MRN_cleaned', 'education','ethnicity',"hh_relationship"  ,        "relationship_duration","pneu_vax_dic" ,"pneu_vax_date" ,"weight", "height","pos_covid_past" ,          "pos_covid_date","smoke_dic", "contact_children" ,       
"age_child_contacts"  ,     "frequency_child_contacts" ,"time_day_child_contacts" )]

baseline.chars <- exp1[,c('ID','Household', 'education','ethnicity',"hh_relationship"  ,        "relationship_duration","pneu_vax_dic" ,"pneu_vax_date" ,"weight", "height","pos_covid_past" ,          "pos_covid_date","smoke_dic")]

baseline.chars <- baseline.chars %>%
  group_by(ID) %>%
  mutate(smoke_dic= max(smoke_dic, na.rm=T))

baseline.chars <- baseline.chars %>%
  group_by(ID) %>%
  mutate(repN=row_number())
baseline.chars <- baseline.chars[baseline.chars$repN==1,]
baseline.chars$repN <- NULL

exp2 <- merge(id.ages, exp1[c('ID','time','piab', 'lyta',"contact_children" ,       
"age_child_contacts"  ,     "frequency_child_contacts" ,"time_day_child_contacts")], by='ID', all.y=T)

exp3 <- merge(exp2,baseline.chars, by='ID', all.x=T )
exp3 <- exp3[order(exp3$ID, exp3$time),]

write.csv(exp3,'./Data/Confidential/export_pfizer.csv')

e3$month <- month(e3$visit_date_cleaned)


e3$contact_children[is.na(e3$contact_children) | e3$contact_children=='NA'] <- 'missing'

e3$child_contact_u12m <- 'No'
e3$child_contact_u12m[grep('<12 m', e3$age_child_contacts)] <- 'Yes'
e3$child_contact_u12m[is.na(e3$age_child_contacts)] <- 'Not answered'

e3$child_contact_u5y <- 'No'
e3$child_contact_u5y[grep('<12 m', e3$age_child_contacts)] <- 'Yes'
e3$child_contact_u5y[grep('13-23 mo', e3$age_child_contacts)] <- 'Yes'
e3$child_contact_u5y[grep('24-59 months', e3$age_child_contacts)] <- 'Yes'
e3$child_contact_u5y[is.na(e3$age_child_contacts) & e3$contact_children!='No'] <- 'Not answered'


e3$child_contact_u12m <- 'No'
e3$child_contact_u12m[grep('<12 m', e3$age_child_contacts)] <- 'Yes'
e3$child_contact_u12m[is.na(e3$age_child_contacts) & e3$contact_children!='No'] <- 'Not answered'

e3$child_contact_13_23m <- 'No'
e3$child_contact_13_23m[grep('13-23 mo', e3$age_child_contacts)] <- 'Yes'
e3$child_contact_13_23m[is.na(e3$age_child_contacts) & e3$contact_children!='No'] <- 'Not answered'

e3$child_contact_24_59m <- 'No'
e3$child_contact_24_59m[grep('24-59 months', e3$age_child_contacts)] <- 'Yes'
e3$child_contact_24_59m[is.na(e3$age_child_contacts) & e3$contact_children!='No'] <- 'Not answered'

e3$child_contact_5_10y <- 'No'
e3$child_contact_5_10y[grep('5-10 years', e3$age_child_contacts)] <- 'Yes'
e3$child_contact_5_10y[is.na(e3$age_child_contacts) & e3$contact_children!='No'] <- 'Not answered'

e3$child_contact_over10y <- 'No'
e3$child_contact_over10y[grep('>10 years', e3$age_child_contacts)] <-'Yes'
e3$child_contact_over10y[is.na(e3$age_child_contacts) & e3$contact_children!='No'] <- 'Not answered'






```

Have you had COVID?
```{r}

table(e3$time, e3$pos_covid_past)

unique(e3$ID[e3$pos_covid_past=='Yes, it was positive'])
table(e3$pos_covid_date)



```


Do you usually have contact with children?

```{r, echo=F}
table1(~piab_pos|`contact_children` , data= e3 )

fisher.test(e3$piab_pos, e3$contact_children)

```

Contact \<5 y olds

```{r, echo=F}
table1(~piab_pos|`child_contact_u5y` , data= e3 )

fisher.test(e3$piab_pos, e3$child_contact_u5y)

```

How often do you have contact with children?

```{r, echo=F}
table1(~piab_pos|frequency_child_contacts , data= e3 )

```

How much contact per day do you have?

```{r, echo=F}
table1(~piab_pos|`time_day_child_contacts` , data= e3 )

```

What ages do you have contact with?

<12m?

```{r, echo=F}
table1(~piab_pos|child_contact_u12m  , data= e3 )
fisher.test(e3$piab_pos, e3$child_contact_u12m)
```

12-59m?

```{r, echo=F}

e3$child_contact_13_59m <- e3$child_contact_13_23m
e3$child_contact_13_59m[e3$child_contact_24_59m=='Yes'] <- 'Yes'


table1(~piab_pos|child_contact_13_59m  , data= e3 )
fisher.test(e3$piab_pos, e3$child_contact_13_59m)

```

12-23m?

```{r, echo=F}
table1(~piab_pos|child_contact_13_23m  , data= e3 )
fisher.test(e3$piab_pos, e3$child_contact_13_23m)

```

2-5y

```{r, echo=F}
table1(~piab_pos|child_contact_24_59m  , data= e3 )
fisher.test(e3$piab_pos, e3$child_contact_24_59m)

```

5-10y

```{r, echo=F}
table1(~piab_pos|child_contact_5_10y  , data= e3 )

fisher.test(e3$piab_pos, e3$child_contact_5_10y)
```

more than 10y

```{r, echo=F}
table1(~piab_pos|child_contact_over10y   , data= e3 )
fisher.test(e3$piab_pos, e3$child_contact_over10y)

```


2-10years

```{r}
e3$child_contact_2_10y <- NA
e3$child_contact_2_10y[e3$child_contact_24_59m=='No' & e3$child_contact_5_10y=='No'] <- 'No' 
e3$child_contact_2_10y[e3$child_contact_24_59m=='Yes' | e3$child_contact_5_10y=='Yes'] <- 'Yes' 
e3$child_contact_2_10y[e3$child_contact_24_59m=='Not answered' & e3$child_contact_5_10y=='Not answered'] <- 'Not answered'


table1(~piab_pos|child_contact_2_10y   , data= e3 )
fisher.test(e3$piab_pos, e3$child_contact_2_10y)
```

Have you taken part in any social activities or outings during the past two weeks?

```{r, echo=F}
#Need to fix this so that those not doing activities also represented

table1(~piab_pos|recent_social_activities , data= e3 )

```

What sorts of activities have you participated in? - Selected Choice

```{r,echo=F}
#Need to fix this so that those not doing activities also represented

table1(~piab_pos|social_activity_type , data= e3 )

#do it person-wise


e3$soc_act_bin <- 0
e3$soc_act_bin[e3$recent_social_activities=='No'] <- 0
e3$soc_act_bin[e3$recent_social_activities=='Yes'] <- 1

e3$soc_act_family_bin <- 0
e3$soc_act_family_bin[e3$social_activity_type=='Activities with family'] <- 1


e3$soc_act_friend_bin <- 0
e3$soc_act_friend_bin[e3$social_activity_type=='Activities with friends'] <- 1


e3$soc_act_comm_ctr_bin <- 0
e3$soc_act_comm_ctr_bin[e3$social_activity_type=='Activities at community centers'] <- 1

e3$soc_act_fitness_bin <- 0
e3$soc_act_fitness_bin[e3$social_activity_type=='Fitness activities'] <- 1

actitivity_sumamry <- e3 %>%
  group_by(ID) %>%
  summarize(soc_activity= max(soc_act_bin, na.rm=T) ,
            soc_activity_family= max(soc_act_family_bin, na.rm=T) ,
            soc_activity_friend= max(soc_act_friend_bin, na.rm=T) ,
            soc_activity_comm_ctr= max(soc_act_comm_ctr_bin, na.rm=T) ,
            soc_activity_fitness= max(soc_act_fitness_bin, na.rm=T) 
            ) %>%
  ungroup() %>%
  summarize(freq_activity=mean(soc_activity, na.rm=T), N_activity=sum(soc_activity, na.rm=T),
            freq_family=mean(soc_activity_family, na.rm=T),N_activity_family=sum(soc_activity_family, na.rm=T),
            freq_friend=mean(soc_activity_friend, na.rm=T),N_activity_friends=sum(soc_activity_friend, na.rm=T),
            freq_comm_ctr=mean(soc_activity_comm_ctr, na.rm=T),N_activity_comm_ctr=sum(soc_activity_comm_ctr, na.rm=T),    
            freq_fitness=mean(soc_activity_fitness, na.rm=T),N_activity_fitness=sum(soc_activity_fitness, na.rm=T)    
                             
            )
  
actitivity_sumamry

```

## Current Symptoms

Cough?

```{r, echo=F}
table1(~piab_pos|current_cough, data=e3 )
```

Runny Nose?

```{r, echo=F}
table1(~piab_pos|current_runny_nose, data= e3 )
```

Fever?

```{r, echo=F}
table1(~piab_pos|current_fever, data= e3 )
```

Sore throat?

```{r, echo=F}
table1(~piab_pos|current_sore_throat, data= e3 )
```

Nasal?

```{r, echo=F}
table1(~piab_pos|current_nasal, data= e3 )
```

## Pneumooccal vaccination

```{r, echo=F}
table1(~piab_pos|pneu_vax, data= e3 )
```

##COVID History


```{r, echo=F}
table1(~piab_pos|pos_covid_past, data= e3 )

```

## Ethnicity

```{r}
table1(~piab_pos|ethnicity, data= e3 )
```

## Education level

```{r, echo=F}
table1(~education, data= e3[e3$time==1,] ) #89/98 report some college
```

```{r, echo=F}
table1(~ethnicity, data= e3[e3$time==1,] )
```

##Medications Immune meds

```{r, echo=F}
table1(~piab_pos|immuno_meds, data= e3 )
```

Asthma meds

```{r, echo=F}
table1(~piab_pos|asthma_meds, data= e3 )
```

Recent Antibiotics

```{r, echo=F}
table1(~piab_pos|recent_abx , data= e3 )

```

# Seasonality

```{r, echo=F}
table1(~piab_pos|month , data= e3 )
```

## Recent Vaccine receipt (note many missing)

```{r, echo=F}
table1(~piab_pos|recent_vaccines, data= e3 )

```

Simple model

```{r}
mod1 <- glm(piab_pos ~ child_contact_u12m+ child_contact_13_23m + child_contact_24_59m + child_contact_5_10y + child_contact_over10y, data=e3[e3$child_contact_13_23m!='Not answered',], family=binomial(link='log'))

summary(mod1)
```

GEE to account for within-individual repeated sampling
```{r}
mod.df <- e3[e3$child_contact_13_23m!='Not answered',]

mod1.gee <- gee(piab_pos ~ child_contact_u12m+ child_contact_13_23m + child_contact_24_59m + child_contact_5_10y + child_contact_over10y,id=mod.df$ID, data=mod.df, family=binomial(link='log'), corstr="exchangeable")

summary(mod1.gee)

coefs <-mod1.gee$coefficients
coef.lcl <- coefs - 1.96 * sqrt(diag(mod1.gee$naive.variance))
coef.ucl <- coefs + 1.96 * sqrt(diag(mod1.gee$naive.variance))

plot(1:5,coefs[-1])
arrows(1:5, y0=coef.lcl[-1], y1=coef.ucl[-1])
abline(h=0)

```


```{r}
mod1 <- glm(piab_pos ~  child_contact_24_59m , data=e3[e3$child_contact_13_23m!='Not answered',], family=binomial(link='log'))
summary(mod1)
```



There are some survey results missing. We lose information for these people. But we know their child contacts at other time points so can infer

```{r}
e3_filled <- e3 %>%
  group_by(ID) %>%
  mutate(any_contact_u12m= max(child_contact_u12m=='Yes', na.rm=T),
         any_contact_13_23m= max(child_contact_13_23m=='Yes', na.rm=T),
         any_contact_24_59m= max(child_contact_24_59m=='Yes', na.rm=T),
         any_contact_5_10y= max(child_contact_5_10y=='Yes', na.rm=T),
         any_contact_over_10y= max(child_contact_over10y=='Yes', na.rm=T),
         )

e3_filled$any_contact13_59m <- e3_filled$any_contact_13_23m + e3_filled$any_contact_24_59m
e3_filled$any_contact13_59m[e3_filled$any_contact13_59m >1] <- 1

mod2 <- glm(piab_pos ~ any_contact_u12m+ any_contact_13_23m + any_contact_24_59m + any_contact_5_10y + any_contact_over_10y 
            , data=e3_filled, family=binomial(link='logit'))
summary(mod2)



```


## Longitudinal analysis
Uses a multi-state model framework to capture transition probabilities
```{r}

l1 <- e3
l1$day <- 0
l1$day[l1$time==2] <- 14
l1$day[l1$time==3] <- 14*2
l1$day[l1$time==4] <- 14*3
l1$day[l1$time==5] <- 14*4
l1$day[l1$time==6] <- 14*5

l1$state <- NA
l1$state[l1$piab_pos==0 & !is.na(l1$piab_pos)] <- 1
l1$state[l1$piab_pos==1 & !is.na(l1$piab_pos)] <- 2
#l1$state <- as.factor(l1$state)


q1 <- rbind(c( 0,0.01),
           c(0.01,0)
)

l1 <- l1[order(l1$ID, l1$day),]

l2 <- l1[, c('state','child_contact_u5y','child_contact_u12m','child_contact_13_23m',"child_contact_24_59m","child_contact_5_10y",'child_contact_over10y', 'ID','day')]
l2$child_contactu5 =0 #neg or missing
l2$child_contactu5[l2$child_contact_u5y=='Yes'] =1 #neg or missing

l2$child_1_10 <- 0

l2$child_1_10[l2$child_contact_13_23m=='Yes' | l2$child_contact_24_59m=='Yes' | l2$child_contact_5_10y=='Yes'] <- 1

l2$child_2_10 <- 0
l2$child_2_10[ l2$child_contact_24_59m=='Yes' | l2$child_contact_5_10y=='Yes'] <- 1

l2$child_5_10 <- 0
l2$child_5_10[  l2$child_contact_5_10y=='Yes'] <- 1

msm.mod1 <- msm( state~ day ,  subject=ID , qmatrix=q1 , data=l2)
msm.mod1

msm.mod1a <- msm( state~ day ,  subject=ID , qmatrix=q1 , data=l2,covariates = list("1-2" = ~ child_contact_u12m))
msm.mod1a

msm.mod2 <- msm( state~ day ,  subject=ID , qmatrix=q1 , data=l2,covariates = list("1-2" = ~ child_1_10))
msm.mod2

msm.mod3 <- msm( state~ day ,  subject=ID , qmatrix=q1 , data=l2,covariates = list("1-2" = ~ child_2_10))
msm.mod3

msm.mod4 <- msm( state ~ day ,  subject=ID , qmatrix=q1 , data=l2,covariates = list("1-2" = ~ child_5_10 ))
msm.mod4

msm.mod5 <- msm( state ~ day ,  subject=ID , qmatrix=q1 , data=l2,covariates = list("1-2" = ~ child_contact_over10y ))
msm.mod5


```

