---
title: "format_data"
author: "Dan Weinberger"
date: "5/25/2021"
output:
  pdf_document: default
  html_document: default
---
## INSTRUCTIONS TO GET SURVEY DATA
1. Go to Yale qualtrics site 
2. Navigate to Intake questionaire.
3. Download data as .xlsx
4. Repeat for fortnightly survey


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readxl)
library(reshape2)
library(dplyr)
library(stringr)

library(pheatmap)
library(RColorBrewer)
library(cowplot)
library(gridExtra)
library(ggplotify)
library(table1)
library(lubridate)
library(patchwork)
source('heat.fun.R')

```


ID 6.1 Saliva was very low under 100 ul.

ID 5.6 qpcr LytA -piaB was ran in plate A155_scope_2021_06_02 .The issue was the label: FIXED 5.5 to 5.6

ID 12.4 qPCR LytA -piaB was ran in plate A150_scope_2021_05_24.The issue was the label --FIXED 12.3 to 12.4

ID 62.2,63.2 samples were picked up 1 month after .We didn’t see any growth after incubation 37 degree over night 

ID 20.5 - 21.5 to verify


```{r, echo=F}
 key1 <- read.csv('./Data/hh_id_key.csv')

master1 <- read_excel('./Data/confidential/SCOPE_Master_20210108.xlsx', sheet='COM SAL')


key2 <- read_excel('./Data/confidential/SCOPE Participant MRN and ID.xlsx', skip=1)
key3 <- cbind.data.frame('id'=c(100:101) , 'Household'=c(51,51) )


key1 <- bind_rows(key1,key3)

file_list <- list.files(path="./Data/cleaned")
  
 all.res <-
   lapply(file_list, function(x){
     print(x)
     ds <- read_excel(paste0("./Data/cleaned/",x), skip=19)
     ds$expt <- substr(x,1,4)
     ds$Sample <- as.character(ds$Sample)
     ds$Target <- as.character(ds$Target)
     ds$Cq <- as.numeric(as.character(ds$Cq))
     cols.keep <- 
       c('Cq','Sample','Target','Content','expt','Biological Set Name')
     ds.names <- names(ds)
     ds.names <- intersect(ds.names,cols.keep)
    ds <- ds[,ds.names]
     return(ds)
   })
 

all.res <- lapply(1:length(all.res), function(x){ all.res[[x]]$plate= x 
return(all.res[[x]])
}) 


test1 <- lapply(all.res, function(x)  is.character(x$Target))

b1 <- bind_rows(all.res) 

b1$Target[is.na(b1$Target)] <- 
  b1$`Biological Set Name`[is.na(b1$Target)]

b1 <- b1[b1$Content!='Std',]

b1$Target <- gsub(' ' ,'', b1$Target, fixed=T)

b1$Target <- tolower(b1$Target)

b1$Target[b1$Target=='lyta_sso'] <- 'lyta'
b1$Target[b1$Target=='piab_sso'] <- 'piab'


#b1$Sample[is.na(b1$Sample)] <- b1$`sample id`[is.na(b1$Sample)]

b1 <-b1[!is.na(b1$Sample),]

b1$Sample <- round(as.numeric(b1$Sample),1)

b1 <- b1[b1$Sample>0 & !is.na(b1$Sample),]

c1 <- b1[, c('Sample','Cq','Target','plate')]

c1$Cq[is.na(c1$Cq)] <- 45

c1.m <- reshape2::melt(c1, id.vars= c('Sample','Target','plate'))

c1.c <- reshape2::dcast(c1.m, Sample+plate +variable ~ Target, fun.aggregate = min, fill=999)

c1.c$pos <- c1.c$lyta<40 & c1.c$piab<40

write.csv(c1.c, './Data/cleaned_file.csv')



```

```{r, fig.width=5, fig.height=5}
plot(c1.c$lyta, c1.c$piab, xlim= c(45,0), ylim=c(45,0), xlab='lytA', ylab='piaB', bty='l')
abline(a=0, b=1)
abline(h=40, v=40, col='gray', lty=2)
```


```{r, echo=F}

c1.m$ID <- sub("\\.[0-9]+$", "", c1.m$Sample)
c1.m$ID <- as.numeric(as.character(c1.m$ID))
c1.m$time <- sub('.*\\.', '', c1.m$Sample)

c1.m <- c1.m[order(c1.m$ID, c1.m$time),]

c1.m <- merge(c1.m, key1, by.x='ID',by.y='id', all.x=T)

d1 <- dcast(c1.m[,c('ID','Target','time','value')],  ID +Target ~ time, fun.aggregate = min, na.rm=T, fill=999)

write.csv(d1, './Data/Result1.csv')

d1.alt <- dcast(c1.m[,c('ID','Target','time','value')],  ID +time~ Target, fun.aggregate = min, na.rm=T)

d1.alt <- d1.alt[!(d1.alt$ID %in% c(38,39)),] #these IDs were not included in study

write.csv(d1.alt, './Data/Result2.csv') #Anne's preferred format

c1.m <- c1.m[!(c1.m$ID %in% c(38,39)),] #these IDs were not included in study

d1.a <- acast(c1.m[,c('ID','Target','time','value','Household')], Target ~ ID +Household ~ time, fun.aggregate=min, na.rm=T, fill=9999)

d1.ds <- reshape2::dcast(c1.m[,c('ID','Target','time','value','Household')],   Household +ID+ time ~ Target, fun.aggregate=min, na.rm=T, fill=9999)


```

## Heatmaps
piaB
```{r, fig.width=4, fig.height=8}
piab.plot <- heat.fun(target='piab')
```

lytA
```{r, fig.width=4, fig.height=8}

lyta.plot <- heat.fun(target='lyta')

```

## Summary stats
```{r}
c1.m.m <- melt(c1.m[,c('Target','ID','Sample','value','Household') ],
                     id.vars=c('Target','ID','Sample','Household')) 
                     
c1.m.c <- dcast(c1.m.m, ID + Household + Sample  ~Target, fun.aggregate = min, fill=999)

#N people
length(unique(c1.m.c$ID))

#N households
length(unique(c1.m.c$Household))


#N samples
nrow(c1.m.c)
#length(unique(c1.m.c$Sample[c1.m$Target=='piab']))
#length(unique(c1.m.c$Sample[c1.m$Target=='lyta']))

#N positive
sum(c1.m.c$piab<40 )

#% positive
sum(c1.m.c$piab<40 )/nrow(c1.m.c)

#N positive lytA
sum(c1.m.c$lyta<40 )

sum(c1.m.c$lyta<40 )/nrow(c1.m.c)


#Count N people with piaB positives by IDs
c1.m.spl.id <- split(c1.m.c, c1.m.c$ID)
sum(sapply(c1.m.spl.id, function(x){
  min(x$piab) <40 
}
  ))

#Count N HH with piab positives by IDs
c1.m.spl.id <- split(c1.m.c, c1.m.c$Household)
sum(sapply(c1.m.spl.id, function(x){
  min(x$piab) <40 
}
  ))


```


## Merge in results with master file
--To do:
-harmonize column names between fortnightly and initial questionaire so we can merge
-Fix dates
-Check missing visitNs
```{r, echo=F}

#Intake questionaire
q0 <-  read_excel('./Data/confidential/SCOPE Intake questionnaire_September 24, 2021_11.57.xlsx', skip=1)

select.colnames <- c('MRN', 
            "SCOPE ID (eg. 0007, 0044, 0072 etc) if known/listed", 
            "Date of Visit","Visit Location - Selected Choice",
            "What is your date of birth? (DD-MM-YYYY)", 
            "1.     Are you experiencing any of the following symptoms today?   - Q8#1 - Nasal congestion",
            "1.     Are you experiencing any of the following symptoms today?   - Q8#1 - Coughing",
            "1.     Are you experiencing any of the following symptoms today?   - Q8#1 - Running nose",
            "1.     Are you experiencing any of the following symptoms today?   - Q8#1 - Sore throat", 
            "1.     Are you experiencing any of the following symptoms today?   - Q8#1 - Fever", 
            "1.    \nHave you had a positive test for COVID?",
            "If\nyes, when did you have this positive test? (DD-MM-YYYY)",
            "Have\nyou received the pneumonia vaccine?"  , 
            "If yes,\nwhen did you receive this vaccine? (DD-MM-YYYY)",
           "Have\nyou been hospitalized for pneumonia in the last month (30 days)?" , 
           "Have\nyou taken any antibiotics in the last month (30 days)?",  
           "Weight",
           "Height (feet, inches)","
           Gender - Selected Choice","What is your highest level of education?", 
           "How would you describe your ethnicity? - Selected Choice"  ,
           "What is the relationship between yourself and the person you are enrolling into the study with?", 
           "How long have you been living together? - Years" ,
           "How much time do you spend together each day? - ___ hours", 
           "Have you received the flu shot this season?",
           "Have you had any of the following symptoms in the past 2 weeks? - Q39#1 - Nasal congestion" ,
          "Have you had any of the following symptoms in the past 2 weeks? - Q39#1 - Coughing",
         "Have you had any of the following symptoms in the past 2 weeks? - Q39#1 - Running nose",
         "Have you had any of the following symptoms in the past 2 weeks? - Q39#1 - Sore throat" ,
         "Have you had any of the following symptoms in the past 2 weeks? - Q39#1 - Fever",
         "Have you taken any medications for or been told by a health care provider that you have any of th... - Q40#1 - 1.\tImmunodeficiency (Examples include: HIV/AIDS, corticosteroid medication like “Prednisone”, medicine or therapy to treat a cancerous tumor or mass, have had your spleen removed or have chronic kidney disease)",
         "Have you taken any medications for or been told by a health care provider that you have any of th... - Q40#1 - 2.\tDiabetes",
         "Have you taken any medications for or been told by a health care provider that you have any of th... - Q40#1 - 5c. Chronic lung disease: Asthma",
          "Do you currently smoke cigarettes or e-cigarrettes?",
         "Do you have any regular contact with children?",
         "What is the age range of the children you have had contact with? - Selected Choice" ,
         "How often do you usually have contact with children?" ,
         "How much contact per day do you have?",
         "Time of sample collection:" 
         )

q0 <- q0[,c(which(names(q0) %in% select.colnames),grep('Are you experiencing any of the following symptoms today', names(q0)) )]

names(q0) <- c('MRN','scope_id', 'visit_date','visit_location', 'dob','pos_covid_past', 'pos_covid_date', 'pneu_vax', 'pneu_vax_date','pneumonia_hosp', 'recent_abx', 'weight','height','education','ethnicity', 'hh_relationship', 'relationship_duration','hh_contact_time','flu_shot', 'recent_nasal', 'recent_cough', 'recent_runny_nose', 'recent_sore_throat', 'recent_fever','immuno_meds','diabetes_meds','asthma_meds', 'current_smoke', 'contact_children','age_child_contacts', 'frequency_child_contacts', 'time_day_child_contacts', 'time_sample', 'current_nasal','current_cough', 'current_runny_nose', 'current_sore_throat','current_fever', 'other_symp1','othr_symp2')

q0$visit_date[q0$visit_date %in% c( '12/APR/2021','12 Apr 2021','12/14/2021')] <- '2021-04-12'

q0$visit_date_clean <- parse_date_time(q0$visit_date, orders=c('d-b-Y','d/b/Y','d-m-Y','m/d/Y','m/d/y','Y-m-d'))
q0$visitN <- '1'



#This is fortnightly questionnaire
q1 <- read_excel('./Data/confidential/SCOPE Fortnightly Questionnaire_September 24, 2021_12.00_clean.xlsx', skip=1)

select.colnames.q1 <- c('MRN', "Fortnightly visit number:", "SCOPE ID if known/listed (eg. 0002, 0027, 0044, 0072...)", "Date of Visit", "Visit Location - Selected Choice","Have you taken part in any social activities or outings during the past two weeks?", "What sorts of activities have you participated in? - Selected Choice" , "Have you had any contact with children in the past two weeks?" , "If yes, what is the age range of the children you have had contact with? - Selected Choice","How often do you usually have contact with children?" , "How much contact per day do you have?", "Have you been tested for COVID in the last two weeks?" , "Have you had any sick visits to the doctor or been hospitalized in the last two weeks?" , "Have you received any new vaccines in the last two weeks? - Selected Choice"   ,"Time of sample collection:" ,'Have you taken any antibiotics in the last two weeks?' )

q1 <- q1[,c(which(names(q1) %in% select.colnames.q1),grep('Which symptoms are you experiencing today', names(q1)),grep('Have you had any of the following symptoms', names(q1))  )]

names(q1) <-c('MRN','scope_id', 'visit_date', 'visitN', 'visit_location','recent_social_activities','social_activity_type', 'contact_children','age_child_contacts', 'frequency_child_contacts', 'time_day_child_contacts','recent_covid_test', 'recent_doctor','recent_abx', 'recent_vaccines', 'sample_time', 'current_nasal','current_cough', 'current_runny_nose', 'current_sore_throat','current_fever', 'other_symp1','othr_symp2','recent_nasal', 'recent_cough', 'recent_runny_nose', 'recent_sore_throat', 'recent_fever' , 'recent_other_symp1','recent_other_symp2')
   
q1$visit_date_clean <- parse_date_time(q1$visit_date, orders=c('d-b-Y','d/b/Y','d-m-Y','d-m-y','m/d/Y','m/d/y','Y-m-d'))
           
#which variables are characteristics of the person (not time varying)
q0.baseline.vars <- c('dob','scope_id',"pos_covid_past","pos_covid_date","pneu_vax", "pneu_vax_date",'ethnicity','weight','height','education', "immuno_meds" ,"diabetes_meds", "asthma_meds","flu_shot","hh_relationship", "relationship_duration")

q1 <- bind_rows(q0[,-which(names(q0) %in% q0.baseline.vars)], q1)


q1$MRN <-  gsub("_visit#2.pdf","", q1$MRN )
q1$MRN <-  gsub("MR","", q1$MRN )
q1$MRN <-  gsub("mr","", q1$MRN )
q1$MRN <-  as.numeric(q1$MRN)


q1$visitN <- gsub('V','', q1$visitN)
q1$visitN <- gsub('v','', q1$visitN)

#q1 <- merge(q1, q0[,c('MRN',q0.baseline.vars)], by='MRN', all=T)

write.csv(q1, './data/confidential/survey_data_to_clean.csv')


q1 <- read.csv('./data/confidential/survey_data_CLEANED.csv', sep=';', skip=1)

q1 <- q1[,-which(names(q1) %in% q0.baseline.vars)] #baseline vars are messed up in merge


q0$MRN <-  gsub("_visit#2.pdf","", q0$MRN )
q0$MRN <-  gsub("MR","", q0$MRN )
q0$MRN <-  gsub("mr","", q0$MRN )
q0$MRN <-  as.numeric(q0$MRN)

q1$MRN <-  gsub("MR","", q1$MRN )

merge_key <- unique(q1[, c('MRN',  "MRN..Clean." )])

q0 <- merge(q0, merge_key, by='MRN')

q2 <- merge(q1, q0[,which(names(q0) %in% c('MRN..Clean.',q0.baseline.vars))], by= "MRN..Clean.", all.x=T)

q2 <- q2[q2$visitN %in% c(1,2,3,4,5,6),]

q2$Visit_date..Clean. <- as.Date(q2$Visit_date..Clean., '%m/%d/%Y')

q2$dob_clean <-  parse_date_time(q2$dob, orders=c('%m/%d/%Y', '%m/%d/%y', '%d/%m/%y','%d/%m/%y','%m-%d-%Y','%d-%m-%Y'))
q2$dob_clean[q2$dob=='1945'] <- as.Date('1945-01-01')
q2$dob_clean[q2$dob=='/1947'] <- as.Date('1947-01-01')

q2$dob_clean <- as.Date(q2$dob_clean)

q2$dob_clean <- q2$dob_clean - 100*365*(q2$dob_clean>=as.Date('2000-01-01')) #if DOB >year 2000, indicates problem with date

q2$Visit_date..Clean.[q2$Visit_date..Clean.==as.Date('2012-03-30')] <- as.Date('2021-03-30')

q2$age <- round(as.numeric(q2$Visit_date..Clean. - as.Date(q2$dob_clean))/365) #only recorded at t=1

#View(q2[, c('visitN','dob_clean',"MRN..Clean.",'Visit_date..Clean.') ])

key2$mr_clean <- gsub('MR', '',key2$`MR#`,)

q2 <- merge(q2, key2, by.x="MRN..Clean.", by.y= "mr_clean" , all.x=T)


#View(q2[is.na(q2$age),])

miss.dob <- q0[q0$dob %in% c('N/A',"19-04-2021") | is.na(q0$dob), c('dob','MRN..Clean.','MRN') ]
write.csv(miss.dob,'./Data/confidential/check_dob.csv')
```


```{r, echo=F}
hist(q2$age, main='Distribution of ages', xlab='Age(y)')
```


```{r, echo=F}



# master1$merge_id <- as.numeric(as.character(gsub('CS','',master1$dw_id_full)))
# 
# e1 <- merge(c1.c, master1, by.x='Sample', by.y='merge_id', all.x)
# 
# excel_date_diff <- as.numeric(as.Date("1970-01-01") - as.Date('1899-12-30')) #origin for excel and R is different
# 
# e1$`sample date`[e1$`sample date`=='2/15/25021'] <- as.character(as.numeric(as.Date('2021-02-15')) + excel_date_diff )
# 
# 
# 
# e1$pos <- 1*(e1$piab<40)
# 
# e1$sample_Date <- as.Date(as.numeric(e1$`sample date`), origin=as.Date('1899-12-30')) #use excel origin of 12-30-1899
# 
# hist(e1$sample_Date, breaks=10)
# e1$part_id <- as.numeric(e1$number)
# 
# e2 <- merge(e1, key2, by.x='part_id', by.y="ID Number", all=T)
# e2$`MR#` <- as.numeric(gsub("MR","", e2$`MR#`  ))


##NEED TO FIX THIS MERGE--MR and visitN could be missing
e3 <- merge(d1.ds, q2, by.x=c('ID','time'), by.y=c("ID Number",'visitN'), all=T)  


e3$social_activity_type <- as.character(e3$social_activity_type)
e3$social_activity_type[e3$recent_social_activities=='No'] <- 'No'

e3$frequency_child_contacts <- as.character(e3$frequency_child_contacts)
e3$frequency_child_contacts[e3$contact_children=='No'] <- 'No'

e3$piab_pos <- as.factor(1*(e3$piab <40 & !is.na(e3$piab)))


## export

exp1 <- e3[,c('ID','time', 'age','education','ethnicity',"hh_relationship"  ,        "relationship_duration","pneu_vax" ,"pneu_vax_date" ,"weight", "height","pos_covid_past" ,          "pos_covid_date","current_smoke", "contact_children" ,       
"age_child_contacts"  ,     "frequency_child_contacts" ,"time_day_child_contacts" )]

write.csv(exp1,'./Data/Confidential/export_pfizer.csv')

e3$month <- month(e3$Visit_date..Clean.)

e3$child_contact_u12m <- 'No'
e3$child_contact_u12m[grep('<12 m', e3$age_child_contacts)] <- 'Yes'

e3$child_contact_13_23m <- 'No'
e3$child_contact_13_23m[grep('13-23 mo', e3$age_child_contacts)] <- 'Yes'

e3$child_contact_24_59m <- 'No'
e3$child_contact_24_59m[grep('24-59 months', e3$age_child_contacts)] <- 'Yes'

e3$child_contact_5_10y <- 'No'
e3$child_contact_5_10y[grep('5-10 years', e3$age_child_contacts)] <- 'Yes'

e3$child_contact_over10y <- 'No'
e3$child_contact_over10y[grep('>10 years', e3$age_child_contacts)] <-'Yes'


```

Do you usually have contact with children?

```{r, echo=F}
table1(~piab_pos|`contact_children` , data= e3 )
```

How often do you have contact with children?

```{r, echo=F}
table1(~piab_pos|frequency_child_contacts , data= e3 )

```

How much contact per day do you have?


```{r, echo=F}
table1(~piab_pos|`time_day_child_contacts` , data= e3 )

```


What ages do you have contact with?

<12m?

```{r, echo=F}
table1(~piab_pos|child_contact_u12m  , data= e3 )
```

12-23m?

```{r, echo=F}
table1(~piab_pos|child_contact_13_23m  , data= e3 )
```

2-5y

```{r, echo=F}
table1(~piab_pos|child_contact_24_59m  , data= e3 )
```

5-10y

```{r, echo=F}
table1(~piab_pos|child_contact_5_10y  , data= e3 )
```

more than 10y

```{r, echo=F}
table1(~piab_pos|child_contact_over10y   , data= e3 )
```

Have you taken part in any social activities or outings during the past two weeks?

```{r, echo=F}
#Need to fix this so that those not doing activities also represented

table1(~piab_pos|recent_social_activities , data= e3 )

```

What sorts of activities have you participated in? - Selected Choice

```{r,echo=F}
#Need to fix this so that those not doing activities also represented

table1(~piab_pos|social_activity_type , data= e3 )
```



## Current Symptoms

Cough?

```{r, echo=F}
table1(~piab_pos|current_cough, data=e3 )
```

Runny Nose?

```{r, echo=F}
table1(~piab_pos|current_runny_nose, data= e3 )
```

Fever?

```{r, echo=F}
table1(~piab_pos|current_fever, data= e3 )
```

Sore throat?

```{r, echo=F}
table1(~piab_pos|current_sore_throat, data= e3 )
```

Nasal?

```{r, echo=F}
table1(~piab_pos|current_nasal, data= e3 )
```

## Pneumooccal vaccination

```{r, echo=F}
table1(~piab_pos|pneu_vax, data= e3 )
```


##COVID History

```{r, echo=F}
table1(~piab_pos|pos_covid_past, data= e3 )

```

## Ethnicity
```{r}
table1(~piab_pos|ethnicity, data= e3 )
```

## Education level

```{r, echo=F}
#table1(~piab_pos|education, data= e3 )
```

#Medications
Immune meds

```{r, echo=F}
table1(~piab_pos|immuno_meds, data= e3 )
```

Asthma meds

```{r, echo=F}
table1(~piab_pos|asthma_meds, data= e3 )
```

Recent Antibiotics

```{r, echo=F}
table1(~piab_pos|recent_abx , data= e3 )

```

# Seasonality 

```{r, echo=F}
table1(~piab_pos|month , data= e3 )
```

## Recent Vaccine receipt (note many missing)
```{r, echo=F}
table1(~piab_pos|recent_vaccines, data= e3 )

```















