---
title: "format_data"
author: "Dan Weinberger"
date: "5/25/2021"
output:
  pdf_document: default
  html_document: default
---

## INSTRUCTIONS TO GET SURVEY DATA

1.  Go to Yale qualtrics site
2.  Navigate to Intake questionaire.
3.  Download data as .xlsx
4.  Repeat for fortnightly survey

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readxl)
library(reshape2)
library(dplyr)
library(stringr)

library(pheatmap)
library(RColorBrewer)
library(cowplot)
library(gridExtra)
library(ggplot2)
library(ggplotify)
library(table1)
library(lubridate)
library(patchwork)
library(msm)
library(gee)
source('heat.fun.R')
source('heat.fun.contact.R')
source('./R/clean_pcr.R')

source('./R/clean_s1_surveys.R')
source('./R/clean_s2_surveys.R')
```

ID 6.1 Saliva was very low under 100 ul.

ID 5.6 qpcr LytA -piaB was ran in plate A155_scope_2021_06_02 .The issue was the label: FIXED 5.5 to 5.6

ID 62.2,63.2 samples were picked up 1 month after .We didn't see any growth after incubation 37 degree over night

ID 20.5 - 21.5 to verify


What to do if we have borderline positive then negative test:

1) If weak positive for piaB (e.g., 35 Ct or higher) AND at least one of the repeats is \<40, then it is positive

2) If weak positive for piaB (e.g., 35 Ct or higher), AND (2 repeats negative) then NEGATIVE

3) if weak positive for piaB and then negative on 1 repeat performed and negative, then negative

4) If positive, take ave of Ct among those \<40


```{r, echo=F}
clean_pcr <- clean_pcr_fun()
```


```{r, fig.width=5, fig.height=5}
plot(clean_pcr$pcr_summary_long$lyta, clean_pcr$pcr_summary_long$piab, xlim= c(45,0), ylim=c(45,0), xlab='lytA', ylab='piaB', bty='l')
abline(a=0, b=1)
abline(h=40, v=40, col='gray', lty=2)
```

## Heatmaps

piaB

```{r, fig.width=8, fig.height=8}
piab.plot <- heat.fun(target='piab')

# breakslist<-seq(15,45,by=1)
# 
# cols1 <- colorRampPalette(c('red4','red','darkorange','orange','gold','lightgoldenrod1','white','white'))(length(breakslist))
# 
# p1 <- ggplot(d1.ds, aes(time, HH_order, fill= piab)) +
#   geom_tile() +
#   scale_fill_viridis_c(option = "B", direction = -1) +
# 
#   theme_classic()+
#   ylab('Serotype tested')+
#    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
#   facet_wrap(~Household, ncol=3)+
#   theme(
#   strip.text.x = element_blank()
# )
# p1
```

lytA

```{r, fig.width=4, fig.height=8}

lyta.plot <- heat.fun(target='lyta')

```

Contacts

## Summary stats

```{r}
c1.m.m <- melt(clean_pcr$pcr_summary[,c('Target','ID','Sample','ct_pos2','Household') ],
                     id.vars=c('Target','ID','Sample','Household')) 
                     
c1.m.c <- dcast(c1.m.m, ID + Household + Sample  ~Target, fun.aggregate = min, fill=999)

#N people
N.id <- length(unique(c1.m.c$ID))
N.id

#N households
N.hh<- length(unique(c1.m.c$Household))
N.hh

#N samples
nrow(c1.m.c)
#length(unique(c1.m.c$Sample[c1.m$Target=='piab']))
#length(unique(c1.m.c$Sample[c1.m$Target=='lyta']))

#N positive
sum(c1.m.c$piab<40 )

sum(c1.m.c$piab<40  &c1.m.c$lyta>=40)
sum(c1.m.c$piab<40  &c1.m.c$lyta<40)

#% positive
sum(c1.m.c$piab<40 )/nrow(c1.m.c)

#N positive lytA
sum(c1.m.c$lyta<40 )

sum(c1.m.c$lyta<40 )/nrow(c1.m.c)


#Count N people with piaB positives by IDs
c1.m.spl.id <- split(c1.m.c, c1.m.c$ID)
N_pos_people <- sum(sapply(c1.m.spl.id, function(x){
  min(x$piab) <40 
}
  ))
N_pos_people
N_pos_people/N.id

#Count N HH with piab positives by IDs
c1.m.spl.id <- split(c1.m.c, c1.m.c$Household)
n_pos_hh <- sum(sapply(c1.m.spl.id, function(x){
  min(x$piab) <40 
}
  ))
n_pos_hh
n_pos_hh/N.hh

# ## Ave age
# age1 <- read_excel('./Data/confidential/AGE Scope Participants samples pick-up 06-15-2021.xlsx', skip=1)
# names(age1)[1] <- 'ID'
# age1 <- age1[ which(age1$ID %in% unique(c1.m.c$ID)),]
# mean(age1$AGE)
# range(age1$AGE)

```

Season 1 surveys
```{r}

s1_data <- clean_s1_surveys()


```

Season 2 surveys
```{r}
s2_data <- clean_s2_surveys()
```


Combine S1 and S2 data
```{r}

s1 <- s1_data$survey_and_pcr_s1 %>%
  select(ID, time, Household, lyta, piab,activities, activity_family, activity_friends, activity_community_center, acitivity_fitness,child_contact,child_contact_u12m,child_contact_13_23m,child_contact_24_59m ,child_contact_5_10y,child_contact_over10y , Age, Gender, Race, Ethnicity)

s2 <- s2_data$survey_and_pcr_s2  %>%
  select(ID, time, Household, lyta, piab,activities, activity_family, activity_friends, activity_community_center, acitivity_fitness,child_contact, child_contact_u12m,child_contact_13_23m,child_contact_24_59m ,child_contact_5_10y,child_contact_over10y, Age, Gender, Race, Ethnicity )

ds.c <- bind_rows(s1, s2) %>%
  mutate(season= str_extract(ID, "[^_]+"), 
         piab_pos=1*(piab<45))

```

Overall prevalence by season
```{r}

ds.c %>%
  group_by(season) %>%
  summarize(prevalence= mean(piab_pos, na.rm=T), N_pos=sum(piab_pos, na.rm=T), n_test=n())
```

Differences by sex
```{r}
ds.c$Gender[ds.c$Gender=='Female'] <- 'F'
ds.c$Gender[ds.c$Gender=='Male'] <- 'M'

table(ds.c$piab_pos, ds.c$Gender)

ds.c %>%
  filter(piab_pos==1) %>%
  group_by(Gender) %>%
  summarize(mean_ct=mean(piab)) 


piab_men <- ds.c %>%
  filter(piab_pos==1 & Gender=='M')

piab_women <- ds.c %>%
  filter(piab_pos==1 & Gender=='F')
  
  wilcox.test(piab_women$piab, piab_men$piab, alternative = "two.sided")

  
  ds.c %>%
  filter(piab_pos==1) %>%
    ggplot( aes(x=Gender,y=piab)) +
    geom_violin()+ theme_classic()
```

Rate of contacts by season
```{r}
ds.c %>%
  group_by(season) %>%
  filter(child_contact %in% c(0,1)) %>%
  mutate(child_contact_u12m = if_else(child_contact_u12m==9999,NA_real_,child_contact_u12m),
         child_contact_13_23m = if_else(child_contact_13_23m==9999,NA_real_,child_contact_13_23m),
         child_contact_u12m = if_else(child_contact_u12m==9999,NA_real_,child_contact_u12m)
         ,child_contact_24_59m = if_else(child_contact_24_59m==9999,NA_real_,child_contact_24_59m)
         ,child_contact_5_10y = if_else(child_contact_5_10y==9999,NA_real_,child_contact_5_10y)) %>%
  summarize(pct_child_contact= mean(child_contact, na.rm=T),  
            pct_child_contact_u12m= mean(child_contact_u12m, na.rm=T),
            pct_child_contact_u13_23m= mean(child_contact_13_23m, na.rm=T),
            pct_child_contact_24_5m= mean(child_contact_24_59m, na.rm=T),
            pct_child_contact_5_10y= mean(child_contact_5_10y, na.rm=T),
            n_test=n())

```



```{r}
prop.table(table(ds.c$child_contact, ds.c$piab_pos), margin=1)
fisher.test(ds.c$piab_pos, ds.c$child_contact)



prop.table(table(ds.c$child_contact_u12m, ds.c$piab_pos), margin=1)
fisher.test(ds.c$piab_pos, ds.c$child_contact_u12m)


prop.table(table(ds.c$child_contact_13_23m, ds.c$piab_pos), margin=1)
fisher.test(ds.c$piab_pos, ds.c$child_contact_13_23m)

prop.table(table(ds.c$child_contact_24_59m, ds.c$piab_pos), margin=1)
fisher.test(ds.c$piab_pos, ds.c$child_contact_24_59m)

prop.table(table(ds.c$child_contact_5_10y, ds.c$piab_pos), margin=1)
fisher.test(ds.c$piab_pos, ds.c$child_contact_5_10y)

prop.table(table(ds.c$child_contact_over10y, ds.c$piab_pos), margin=1)
fisher.test(ds.c$piab_pos, ds.c$child_contact_over10y)

mod_s2 <- glm(piab_pos ~ child_contact_u12m + child_contact_13_23m + child_contact_24_59m + child_contact_5_10y + child_contact_over10y , family='binomial', data=ds.c )
summary(mod_s2)
```

GEE to account for within-individual repeated sampling
```{r}
mod.df <- ds.c %>%
  filter(child_contact_13_23m!=9999 & !is.na(piab_pos)) %>%
  mutate(ID=as.factor(ID))



mod1.gee <- gee(piab_pos ~ child_contact_u12m+ child_contact_13_23m + child_contact_24_59m + child_contact_5_10y + child_contact_over10y,id=mod.df$ID, data=mod.df, family=binomial(link='log'), corstr="exchangeable")

summary(mod1.gee)

coefs <-mod1.gee$coefficients
coef.lcl <- coefs - 1.96 * sqrt(diag(mod1.gee$naive.variance))
coef.ucl <- coefs + 1.96 * sqrt(diag(mod1.gee$naive.variance))

plot(1:5,coefs[-1])
arrows(1:5, y0=coef.lcl[-1], y1=coef.ucl[-1])
abline(h=0)

```



## Longitudinal analysis
Uses a multi-state model framework to capture transition probabilities
```{r}
mod.df <- ds.c %>%
  filter(child_contact_13_23m!=9999 & !is.na(piab_pos)) %>%
  mutate(ID=as.factor(ID))

mod.df$child_contact_u5 <- 
  mod.df$child_contact_u12m + mod.df$child_contact_13_23m + mod.df$child_contact_24_59m


#N contacts
mod.df$child_contact_1_5 <- mod.df$child_contact_13_23m + mod.df$child_contact_24_59m


l1 <- mod.df
l1$day <- 0
l1$day[l1$time==2] <- 14
l1$day[l1$time==3] <- 14*2
l1$day[l1$time==4] <- 14*3
l1$day[l1$time==5] <- 14*4
l1$day[l1$time==6] <- 14*5

l1$state <- NA
l1$state[l1$piab_pos==0 & !is.na(l1$piab_pos)] <- 1
l1$state[l1$piab_pos==1 & !is.na(l1$piab_pos)] <- 2
#l1$state <- as.factor(l1$state)


q1 <- rbind(c( 0,0.01),
           c(0.01,0)
)

l1 <- l1[order(l1$ID, l1$day),]

l2 <- l1[, c('state','child_contact_u5','child_contact_1_5','child_contact','child_contact_u12m','child_contact_13_23m',"child_contact_24_59m","child_contact_5_10y",'child_contact_over10y', 'ID','day')]

msm.mod1 <- msm( state~ day ,  subject=ID , qmatrix=q1 , data=l2)
msm.mod1

msm.mod1a <- msm( state~ day ,  subject=ID , qmatrix=q1 , data=l2,covariates = list("1-2" = ~ child_contact_u12m))
msm.mod1a

msm.mod1b <- msm( state~ day ,  subject=ID , qmatrix=q1 , data=l2,covariates = list("1-2" = ~ child_contact_1_5))
msm.mod1b


msm.mod1b <- msm( state~ day ,  subject=ID , qmatrix=q1 , data=l2,covariates = list("1-2" = ~ child_contact_u5))
msm.mod1b

```




heat map of S1 contacts


```{r, fig.width=4, fig.height=8}
heat.fun.contact(ds=s1_data$contacts_wide_s1)

heat.fun.contact(ds=s2_data$contacts_wide_s2)
```




```{r, echo=F}
hist(age1$AGE, main='Distribution of ages', xlab='Age(y)')
```


Have you had COVID?
```{r}

table(s1_data$survey_and_pcr_s1$time, s1_data$survey_and_pcr_s1$pos_covid_past)

unique(s1_data$survey_and_pcr_s1$ID[s1_data$survey_and_pcr_s1$pos_covid_past=='Yes, it was positive'])
table(s1_data$survey_and_pcr_s1$pos_covid_date)



```


Do you usually have contact with children?

```{r, echo=F}
table1(~piab_pos|`contact_children` , data= s1_data$survey_and_pcr_s1 )

fisher.test(s1_data$survey_and_pcr_s1$piab_pos, s1_data$survey_and_pcr_s1$contact_children)

```

Contact \<5 y olds

```{r, echo=F}
table1(~piab_pos|`child_contact_u5y` , data= s1_data$survey_and_pcr_s1 )

fisher.test(s1_data$survey_and_pcr_s1$piab_pos, s1_data$survey_and_pcr_s1$child_contact_u5y)

```

How often do you have contact with children?

```{r, echo=F}
table1(~piab_pos|frequency_child_contacts , data= s1_data$survey_and_pcr_s1 )

```

How much contact per day do you have?

```{r, echo=F}
table1(~piab_pos|`time_day_child_contacts` , data= s1_data$survey_and_pcr_s1 )

```

5-10y

```{r, echo=F}
table1(~piab_pos|child_contact_5_10y  , data= s1_data$survey_and_pcr_s1 )

fisher.test(s1_data$survey_and_pcr_s1$piab_pos, s1_data$survey_and_pcr_s1$child_contact_5_10y)
```

more than 10y

```{r, echo=F}
table1(~piab_pos|child_contact_over10y   , data= s1_data$survey_and_pcr_s1 )
fisher.test(s1_data$survey_and_pcr_s1$piab_pos, s1_data$survey_and_pcr_s1$child_contact_over10y)

```


2-10years

```{r}
s1_data$survey_and_pcr_s1$child_contact_2_10y <- NA
s1_data$survey_and_pcr_s1$child_contact_2_10y[s1_data$survey_and_pcr_s1$child_contact_24_59m=='No' & s1_data$survey_and_pcr_s1$child_contact_5_10y=='No'] <- 'No' 
s1_data$survey_and_pcr_s1$child_contact_2_10y[s1_data$survey_and_pcr_s1$child_contact_24_59m=='Yes' | s1_data$survey_and_pcr_s1$child_contact_5_10y=='Yes'] <- 'Yes' 
s1_data$survey_and_pcr_s1$child_contact_2_10y[s1_data$survey_and_pcr_s1$child_contact_24_59m=='Not answered' & s1_data$survey_and_pcr_s1$child_contact_5_10y=='Not answered'] <- 'Not answered'


table1(~piab_pos|child_contact_2_10y   , data= s1_data$survey_and_pcr_s1 )
fisher.test(s1_data$survey_and_pcr_s1$piab_pos, s1_data$survey_and_pcr_s1$child_contact_2_10y)
```

Have you taken part in any social activities or outings during the past two weeks?

```{r, echo=F}
#Need to fix this so that those not doing activities also represented

table1(~piab_pos|recent_social_activities , data= s1_data$survey_and_pcr_s1 )

```

What sorts of activities have you participated in? - Selected Choice

```{r,echo=F}
#Need to fix this so that those not doing activities also represented

table1(~piab_pos|social_activity_type , data= s1_data$survey_and_pcr_s1 )

#do it person-wise

actitivity_sumamry

```

## Current Symptoms

Cough?

```{r, echo=F}
table1(~piab_pos|current_cough, data=s1_data$survey_and_pcr_s1 )
```

Runny Nose?

```{r, echo=F}
table1(~piab_pos|current_runny_nose, data= s1_data$survey_and_pcr_s1 )
```

Fever?

```{r, echo=F}
table1(~piab_pos|current_fever, data= s1_data$survey_and_pcr_s1 )
```

Sore throat?

```{r, echo=F}
table1(~piab_pos|current_sore_throat, data= s1_data$survey_and_pcr_s1 )
```

Nasal?

```{r, echo=F}
table1(~piab_pos|current_nasal, data= s1_data$survey_and_pcr_s1 )
```

## Pneumooccal vaccination

```{r, echo=F}
table1(~piab_pos|pneu_vax, data= s1_data$survey_and_pcr_s1 )
```

##COVID History


```{r, echo=F}
table1(~piab_pos|pos_covid_past, data= s1_data$survey_and_pcr_s1 )

```

## Ethnicity

```{r}
table1(~piab_pos|ethnicity, data= s1_data$survey_and_pcr_s1 )
```

## Education level

```{r, echo=F}
table1(~education, data= s1_data$survey_and_pcr_s1[s1_data$survey_and_pcr_s1$time==1,] ) #89/98 report some college
```

```{r, echo=F}
table1(~ethnicity, data= s1_data$survey_and_pcr_s1[s1_data$survey_and_pcr_s1$time==1,] )
```

##Medications Immune meds

```{r, echo=F}
table1(~piab_pos|immuno_meds, data= s1_data$survey_and_pcr_s1 )
```

Asthma meds

```{r, echo=F}
table1(~piab_pos|asthma_meds, data= s1_data$survey_and_pcr_s1 )
```

Recent Antibiotics

```{r, echo=F}
table1(~piab_pos|recent_abx , data= s1_data$survey_and_pcr_s1 )

```

# Seasonality

```{r, echo=F}
table1(~piab_pos|month , data= s1_data$survey_and_pcr_s1 )
```

## Recent Vaccine receipt (note many missing)

```{r, echo=F}
table1(~piab_pos|recent_vaccines, data= s1_data$survey_and_pcr_s1 )

```

Simple model

```{r}
mod1 <- glm(piab_pos ~ child_contact_u12m+ child_contact_13_23m + child_contact_24_59m + child_contact_5_10y + child_contact_over10y, data=s1_data$survey_and_pcr_s1[s1_data$survey_and_pcr_s1$child_contact_13_23m!='Not answered',], family=binomial(link='log'))

summary(mod1)
```

GEE to account for within-individual repeated sampling
```{r}
mod.df <- s1_data$survey_and_pcr_s1[s1_data$survey_and_pcr_s1$child_contact_13_23m!='Not answered',]

mod1.gee <- gee(piab_pos ~ child_contact_u12m+ child_contact_13_23m + child_contact_24_59m + child_contact_5_10y + child_contact_over10y,id=mod.df$ID, data=mod.df, family=binomial(link='log'), corstr="exchangeable")

summary(mod1.gee)

coefs <-mod1.gee$coefficients
coef.lcl <- coefs - 1.96 * sqrt(diag(mod1.gee$naive.variance))
coef.ucl <- coefs + 1.96 * sqrt(diag(mod1.gee$naive.variance))

plot(1:5,coefs[-1])
arrows(1:5, y0=coef.lcl[-1], y1=coef.ucl[-1])
abline(h=0)

```


```{r}
mod1 <- glm(piab_pos ~  child_contact_24_59m , data=s1_data$survey_and_pcr_s1[s1_data$survey_and_pcr_s1$child_contact_13_23m!='Not answered',], family=binomial(link='log'))
summary(mod1)
```



There are some survey results missing. We lose information for these people. But we know their child contacts at other time points so can infer

```{r}
s1_data$survey_and_pcr_s1_filled <- s1_data$survey_and_pcr_s1 %>%
  group_by(ID) %>%
  mutate(any_contact_u12m= max(child_contact_u12m=='Yes', na.rm=T),
         any_contact_13_23m= max(child_contact_13_23m=='Yes', na.rm=T),
         any_contact_24_59m= max(child_contact_24_59m=='Yes', na.rm=T),
         any_contact_5_10y= max(child_contact_5_10y=='Yes', na.rm=T),
         any_contact_over_10y= max(child_contact_over10y=='Yes', na.rm=T),
         )

s1_data$survey_and_pcr_s1_filled$any_contact13_59m <- s1_data$survey_and_pcr_s1_filled$any_contact_13_23m + s1_data$survey_and_pcr_s1_filled$any_contact_24_59m
s1_data$survey_and_pcr_s1_filled$any_contact13_59m[s1_data$survey_and_pcr_s1_filled$any_contact13_59m >1] <- 1

mod2 <- glm(piab_pos ~ any_contact_u12m+ any_contact_13_23m + any_contact_24_59m + any_contact_5_10y + any_contact_over_10y 
            , data=s1_data$survey_and_pcr_s1_filled, family=binomial(link='logit'))
summary(mod2)



```


## Longitudinal analysis
Uses a multi-state model framework to capture transition probabilities
```{r}

l1 <- s1_data$survey_and_pcr_s1
l1$day <- 0
l1$day[l1$time==2] <- 14
l1$day[l1$time==3] <- 14*2
l1$day[l1$time==4] <- 14*3
l1$day[l1$time==5] <- 14*4
l1$day[l1$time==6] <- 14*5

l1$state <- NA
l1$state[l1$piab_pos==0 & !is.na(l1$piab_pos)] <- 1
l1$state[l1$piab_pos==1 & !is.na(l1$piab_pos)] <- 2
#l1$state <- as.factor(l1$state)


q1 <- rbind(c( 0,0.01),
           c(0.01,0)
)

l1 <- l1[order(l1$ID, l1$day),]

l2 <- l1[, c('state','child_contact_u5y','child_contact_u12m','child_contact_13_23m',"child_contact_24_59m","child_contact_5_10y",'child_contact_over10y', 'ID','day')]
l2$child_contactu5 =0 #neg or missing
l2$child_contactu5[l2$child_contact_u5y=='Yes'] =1 #neg or missing

l2$child_1_10 <- 0

l2$child_1_10[l2$child_contact_13_23m=='Yes' | l2$child_contact_24_59m=='Yes' | l2$child_contact_5_10y=='Yes'] <- 1

l2$child_2_10 <- 0
l2$child_2_10[ l2$child_contact_24_59m=='Yes' | l2$child_contact_5_10y=='Yes'] <- 1

l2$child_5_10 <- 0
l2$child_5_10[  l2$child_contact_5_10y=='Yes'] <- 1

msm.mod1 <- msm( state~ day ,  subject=ID , qmatrix=q1 , data=l2)
msm.mod1

msm.mod1a <- msm( state~ day ,  subject=ID , qmatrix=q1 , data=l2,covariates = list("1-2" = ~ child_contact_u12m))
msm.mod1a

msm.mod2 <- msm( state~ day ,  subject=ID , qmatrix=q1 , data=l2,covariates = list("1-2" = ~ child_1_10))
msm.mod2

msm.mod3 <- msm( state~ day ,  subject=ID , qmatrix=q1 , data=l2,covariates = list("1-2" = ~ child_2_10))
msm.mod3

msm.mod4 <- msm( state ~ day ,  subject=ID , qmatrix=q1 , data=l2,covariates = list("1-2" = ~ child_5_10 ))
msm.mod4

msm.mod5 <- msm( state ~ day ,  subject=ID , qmatrix=q1 , data=l2,covariates = list("1-2" = ~ child_contact_over10y ))
msm.mod5


```

